Merging merges, more or less

### 2015-04-01 2015mergingmerges/section/03-Merging.tex
add example

The previous cases have been rather trivial as they require no synchronisation between the machines, but the remaining cases of @Pull@, @Release@, @Out@ and @OutDone@ are more complicated.

@@

The cases for @OutDone@ proceed similarly to @Out@.

@@

>>>>

The previous cases have been rather trivial as they require no synchronisation between the machines, but the remaining cases of @Pull@, @Release@, @Out@ and @OutFinished@ are more complicated.

@@

The cases for @OutFinished@ proceed similarly to @Out@.

@@

\subsubsection{Example}
Let us merge a @map@ (figure~\ref{fig:com:map}) and a @filter@ (figure~\ref{fig:com:filter}).
The actual program would be
\begin{code}
fun xs
 = let m  = map    (+1) xs
       f  = filter (>0) xs
   in (m, f)
\end{code}

We need to name each state.
While the graphical form is easier to understand, a textual form makes looking up a particular state's transitions easier.
The @map@ state types and transitions are below, with initial state being @m1@.

\begin{code}
m1 : Pull xs     { Some => m2; None => m4 }
m2 : Out m (x+1) { Unit => m3 }
m3 : Release x   { Unit => m1 }
m4 : OutFinished m{ Unit => m5 }
m5 : Done
\end{code}

Likewise, the filter has initial state @f1@.

\begin{code}
f1 : Pull xs     { Some => f2; None  => f5 }
f2 : If (x>0)    { True => f3; False => f4 }
f3 : Out f x     { Unit => f4 }
f4 : Release x   { Unit => f1 }
f5 : OutFinished f{ Unit => f6 }
f6 : Done
\end{code}

We start by computing $@merge@'$ of an empty machine, empty $\psi$ sets, and initial states of each machine:
$$
@merge@'~@m1@~\emptyset~@f1@~\emptyset~@empty@
$$

In this case, there are two possible moves: either @m@ may take a step, or @f@ may take a step, but both steps are @Pull xs@. When two moves are possible, we will take the @m@ one, but this decision has no effect on the outcome \TODO{(prove it)}.
\begin{code}
move m f m1 {} f1 {} =
    Pull xs,
    { Some => m2, {Value xs},    f1, {Value xs}
    ; None => m4, {Finished xs}, f1, {Finished xs} }
(Shared pull, no pending values)
\end{code}

We insert this state and transitions into the empty machine, noting that the transitions to undefined states will be defined by later moves.

\begin{code}
m1,{},f1,{}
    : Pull xs
    { Some => m2, {Value xs},    f1, {Value xs}
    ; None => m4, {Finished xs}, f1, {Finished xs} }
\end{code}

Now, calculate the first undefined move.
Again, either machine may move: @m@ could compute its output, or @f@ could skip past the @Pull xs@, since this has been performed by @m@, witnessed by the @Value xs@ in @f@'s $\psi$ set.

\begin{code}
move m f m2 {Value xs} f1 {Value xs} =
    Out m (x+1)
    { Unit => m3, {Value xs}, f1, {Value xs} }
(Local output, free to write)
\end{code}

This puts @m@ at a @Release@, which is changed to a skip as the later release by @f@ will be the real release.

\begin{code}
move m f m3 {Value xs} f1 {Value xs} =
    Skip
    { Unit => m1, {}, f1, {Value xs} }
(Shared release: first release is skip)
\end{code}

Now machine @m@ is back at a @Pull xs@, but it is unable to execute the pull itself as the other machine @f@ still has a pending @Value xs@ to deal with.
The machine @f@ is also at a @Pull xs@, and is able to skip past since it already has a value.

\begin{code}
move m f m1 {} f1 {Value xs} =
    Skip
    { Unit => m1, {}, f2, {Value xs} }
(Pull has value)
\end{code}

Now @f@ is at an @If@, and @m@ still cannot run.

\begin{code}
move m f m1 {} f2 {Value xs} =
    If (x>0)
    { True  => m1, {}, f3, {Value xs}
    , False => m1, {}, f4, {Value xs} }
(If)
\end{code}

The first case of the @If@ is a local output. 
\begin{code}
move m f m1 {} f3 {Value xs} =
    Out f x
    { Unit => m1, {}, f4, {Value xs} }
(Local output, free to write)
\end{code}

Next is a release.
In this case, the input is shared and the other machine has already released, since there is no @Value xs@ in its set.
We emit an actual release in this case, since we are sure both machines are finished using it.

\begin{code}
move m f m1 {} f4 {Value xs} =
    Release x
    { Unit => m1, {}, f1, {} }
(Shared release: other machine already released)
\end{code}

Now the only missing state is @m4, {Finished xs}, f1, {Finished xs}@ after the @m@'s pull is empty.

\begin{code}
move m f m4 {Finished xs} f1 {Finished xs} =
    OutFinished m
    { Unit => m5, {Finished xs}, f1, {Finished xs} }
(Finish local output)
\end{code}

Now @m@ is @Done@, and cannot move, while @f@ is at a @Pull xs@ that it already knows is finished.
\begin{code}
move m f m5 {Finished xs} f1 {Finished xs} =
    Skip
    { Unit => m5, {Finished xs}, f5, {Finished xs} }
(Pull of finished)
\end{code}

Now @m@ is @Done@, and cannot move, while @f@ is at a @Pull xs@ that it already knows is finished.
\begin{code}
move m f m5 {Finished xs} f5 {Finished xs} =
    OutFinished f
    { Unit => m5, {Finished xs}, f6, {Finished xs} }
(Finish local output)
\end{code}

Both machines are now done.
\begin{code}
move m f m5 {Finished xs} f6 {Finished xs} =
    Done
(Both machines done)
\end{code}

The entire machine at the end is:
\begin{code}
m1,{},f1,{}
    : Pull xs
    { Some => m2, {Value xs},    f1, {Value xs}
    ; None => m4, {Finished xs}, f1, {Finished xs} }
m2,{Value xs},f1,{Value xs}
    : Out m (x+1)
    { Unit => m3, {Value xs},    f1, {Value xs}    }
m3,{Value xs},f1,{Value xs}
    : Skip
    { Unit => m1, {},            f1, {Value xs}    }
m1,{},f1,{Value xs}
    : Skip
    { Unit => m1, {},            f2, {Value xs}    }
m1,{},f2,{Value xs}
    : If (x>0)
    { True => m1, {},            f3, {Value xs}
    , False=> m1, {},            f4, {Value xs}    }
m1,{},f3,{Value xs}
    : Out f x
    { Unit => m1, {},            f4, {Value xs}    }
m1,{},f4,{Value xs}
    : Release x
    { Unit => m1, {},            f1, {}            }
m4,{Finished xs},f1,{Finished xs}
    : OutFinished m
    { Unit => m5, {Finished xs}, f1, {Finished xs} }
m5,{Finished xs},f1,{Finished xs}
    : Skip
    { Unit => m5, {Finished xs}, f5, {Finished xs} }
m5,{Finished xs},f5,{Finished xs}
    : OutFinished f
    { Unit => m5, {Finished xs}, f6, {Finished xs} }
m5,{Finished xs},f6,{Finished xs}
    : Done
\end{code}

The @Skip@s can then be removed quite easily, leading to the graph below.

\begin{dot2tex}[scale=0.5]
digraph {
    rankdir=LR;
    mindist=0.1
    nodesep=0.1;
    ranksep=0.1;
    node [shape="circle", margin=0];
    10 [label="Pull xs"];
    20 [label="Out m (x+1)"];
    50 [label="If"];
    60 [label="Out f x"];
    70 [label="Release x"];
    900[label="OutFinished m"];
    920[label="OutFinished f"];
    930[label="Done"];

    10 -> 20 [label="Some x"];
    10 -> 900 [label="None"];

    50 -> 60 [label="True"];
    50 -> 70 [label="False"];

    20 -> 50;
    60 -> 70;
    70 -> 10;
    900 -> 920;
    920 -> 930;
}
\end{dot2tex}
